Rectangulotopes

### 2024-04-05 draft/rectangulotopes.tex
corrected one ref

\begin{figure}
	\capstart
	\centerline{\includegraphics[scale=1]{shardPolytopes}}
	\caption{The shard polytopes of the up arc~$(1, 4, \{2, 3\}, \varnothing)$, down arc~$(1, 4, \varnothing, \{2, 3\})$, yin arc~$(1, 4, \{2\}, \{3\})$, and yang arc~$(1, 4, \{3\}, \{2\})$. The vertices of the shard polytope of an arc~$\alpha \eqdef (a, b, A, B)$ are labeled by the corresponding $\alpha$-alternating matchings, where we use solid dots~$\bullet$ for elements in~$\{a\} \cup A$ and hollow dots~$\circ$ for elements in~$B \cup \{b\}$. The corresponding vertex coordinates are directly read replacing~$\bullet$ by~$1$ and~$\circ$ by~$-1$. For instance, the vertex labeled~${\bullet \cdot \cdot \,\circ}$ has coordinates~$(1,0,0,-1)$. Adapted from~\cite[Fig.~10]{PadrolPilaudRitter}.}
	\label{fig:shardPolytiopes}
\end{figure}

@@

xGiven a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$.
These posets were characterized (and called Baxter posets) by E.~Meehan~\cite{MR4014603}.
The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
We define the set~$\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
\[
\mathcal{L}(P_w(R)) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
\]

>>>>

\begin{figure}
	\capstart
	\centerline{\includegraphics[scale=1]{shardPolytopes}}
	\caption{The shard polytopes of the up arc~$(1, 4, \{2, 3\}, \varnothing)$, down arc~$(1, 4, \varnothing, \{2, 3\})$, yin arc~$(1, 4, \{2\}, \{3\})$, and yang arc~$(1, 4, \{3\}, \{2\})$. The vertices of the shard polytope of an arc~$\alpha \eqdef (a, b, A, B)$ are labeled by the corresponding $\alpha$-alternating matchings, where we use solid dots~$\bullet$ for elements in~$\{a\} \cup A$ and hollow dots~$\circ$ for elements in~$B \cup \{b\}$. The corresponding vertex coordinates are directly read replacing~$\bullet$ by~$1$ and~$\circ$ by~$-1$. For instance, the vertex labeled~${\bullet \cdot \cdot \,\circ}$ has coordinates~$(1,0,0,-1)$. Adapted from~\cite[Fig.~10]{MR4584712}.}
	\label{fig:shardPolytiopes}
\end{figure}

@@

Given a rectangulation~$R$, the source and target trees~$S(R)$ and~$T(R)$ define two partial orders on $[n]$ (the horizontal edges are oriented from left to right and the vertical edges are oriented from bottom to top).
The \defn{weak poset}~$P_w(R)=([n],\prec_w)$ of the rectangulation~$R$ is the transitive closure of the union of the two partial orders defined by the intertwining binary trees~$S(R)$ and~$T(R)$.
These posets were characterized (and called Baxter posets) by E.~Meehan~\cite{MR4014603}.
The weak poset is a two-dimensional lattice, whose minimum is the root of~$S(R)$ and whose maximum is the root of~$T(R)$.
We define the set~$\mathcal{L}(P_w(R))$ as the linear extensions of the weak poset:
\[
\mathcal{L}(P_w(R)) \eqdef \bigset{\pi\in\f{S}_n }{ i\prec_w j\implies \pi^{-1}(i) < \pi^{-1}(j)}.
\]
